Supplementary Material

### 2025-10-08 manuscript.tex
remaking plots
manuscript edits

Along with the CDF and PDF, the QF is a defining function of a random variable. 
The QF, however, often receives less attention in standard statistical 
training than do the CDF and PDF \cite[]{parzen2004quantile}, and the CDF and 
PDF are more used often for modeling and inference.
% But the QF tends to be the little brother 
% \jarad{little brother?}
% of both the CDF and PDF in that the latter two functions are taught and used far more in statistical modeling and inference. 
% Nevertheless, the QF and several of its properties are central to the QGP model introduced in this paper. 
% \jarad{another throwaway sentence}
This section contains the definition of the QF, important properties, the 
definition of sample quantiles, and key central limit theorems (CLTs) of sample 
quantiles.

@@

The covariance function $\kappa_{\theta} (\cdot, \cdot)$ initially appears as 
if it will be difficult to evaluate and indeed can be, depending on the 
functional forms of $f_{\theta}$ and $Q_{\theta}$. But where these functions 
belong to a location-scale family, the covariance function can be greatly 
simplified. 
Noting that the covariance is a function of $f_{\theta}(Q_{\theta}(p)) = q_{\theta}(p)$, the QDF, if $\theta = (\mu, \sigma)$ where $\mu$ is a location parameter and $\sigma$ is a scale parameter, then $q_{\theta}(p) = \sigma q(p)$ where $q(p)$ is the QDF of a standard distribution having location 0 and shape 1. \cite{staudte2017shapes} calls this the location invariant and scale equivariant property. For location-scale distribution families, this greatly simplifies the form of $\kappa_{\theta}(\cdot,\cdot)$ so that it becomes
\[
    \kappa_{\theta}(p, p') = \sigma^{2}[p\wedge p' - p p'] q(p)q(p') 
\] 
Thus for estimating $\theta$ one only needs to estimate $\sigma^2$. A special 
case of the QGP for QM of a location-scale family is the normal QGP which we 
define and explore below.

@@

In section \ref{sec:simulation_analyses}, we analyze PIT transformed QGP model 
(\ref{eq:pit_qgp}) where $F_{\theta}$ is set as a normal mixture distribution 
with $C = 4$ component normal distributions. Fitting this model proved faster 
and produced better results than trying to fit the QGP model (\ref{eq:qgp}) and 
evaluating $Q_{\theta}$ the QF of a normal mixture. In fact we found that in 
cases where $F_{\theta}$ was a simpler function, such as a normal or 
exponential distribution, fitting model (\ref{eq:pit_qgp}) was at least as fast 
and the results were at least as good as fitting (\ref{eq:qgp}). Thus when 
possible, we elected to fit (\ref{eq:pit_qgp}).

@@

A number of metrics for measuring the distance between two continuous 
univariate distributions exist. Those we consider in this paper are the 
Wasserstein distance (WD) and a slight modification of it, the total variation 
(TV) or the statistical difference, and the Kullback-Leibler divergence (KLD). 
A brief overview of these and other metrics, some of their statistical uses 
and properties, and relationships between metrics is found in 
\cite{gibbs2002choosing}. These metrics were used in section 
\ref{sec:simulation_analyses} to assess the how well different QM methods 
approximate a true distribution.

@@

We define a modified version of the WD, the uniform 1-WD (UWD1) in 
(\ref{eq:pit_wass_dist}). Here we take $F_X$ to be the CDF of a continuous 
random variable $X$. For a different continuous random variable $\xi$ with CDF 
$F_{\xi}$, $F_X(\xi)$ is a random variable with support on $[0,1]$. We let 
$F_{X,\xi}$ be the CDF of the random variable $F_X(\xi)$. Then the measure in 
(\ref{eq:pit_wass_dist}) measures how close $F_X(\xi)$ is the the CDF of 
a standard uniform distribution with CDF $\mathcal{U}$, 
noting that by the PIT, if 
$\xi \overset{D}{=} X$, $F_X(\xi)$ is the random variable of a standard uniform 
distribution. The integral in (\ref{eq:pit_wass_dist}) takes values between
0 and 1/2 \cite[]{645854}, and multiplying by two 
ensures the measure takes values between 0 and 1. This makes the UWD1 more
interpretable than the WD
where a value near 0 suggests $\xi$ and $X$ are ``near'' each other, and a value
near 1 suggests they are far apart.

@@

The top of figure \ref{fig:normal_cov_dists} shows the 90\% coverage of 
the posterior distribution credible intervals as $n$ increases for 
$\mu$, $\sigma$ and $n$ for five models. QGP-n and ORD-n models where $n$ is 
known are included along with QGP, ORD, and IND. 
The 90\% credible intervals of the 
parameters were calculated by computing the $0.05^{th}$ and $0.95^{th}$ 
quantiles from the posterior distribution samples. The coverage percentage is 
calculated as the percentage of the 500 replicates which the true parameter 
value was within the 90\% credible interval.
As expected, the nominal coverage for the QGP and ORD models is better than 
that of the IND model, and for the two models where $n$ is known the coverage 
is better than where it is unknown, particularly for lower values of 
$n$ and $K$. The coverage for the ORD models appears to be slightly better 
than for the QGP models though not by much, and the difference
decrases as $n$ increases. 
The superior performance of the ORD model is possibly because it
provides exact inference 
whereas the QGP model provides asymptotic inference.

@@

With the true distribution family being unknown, we made the modeling decision 
to set $F$ in the QGP model of (\ref{eq:pit_qgp}) to be a mixture of normals 
distribution taking the form of (\ref{eq:norm_mix}). The number of component 
distributions was chosen to be $C = 4$ to provide enough flexibility to 
approximate the different shapes of quantile functions. This decision was 
based on the common claim that any continuous distribution may be approximated 
by a finite mixture of normal distributions 
\cite[]{peel2000finite,nguyen2019approximations,nguyen2020approximation}. This 
claim may be optimistic, but we found that a four component normal mixture 
distribution can approximate the non-normally shaped distributions reasonably 
well. However, any function which meets the technical definition of a CDF would 
be appropriate to use for $F$, see \cite{gasthaus2019probabilistic} for an 
example.

The three distributions from which data were simulated and sample quantiles 
estimated and to which QM methods were fit were the extreme value (EV) 
distribution with location 0 and scale 1 $EV(0,1)$, Laplace (La) with location 
0 and scale 1 $La(0,1)$, and a two component normal mixture (MIX) 
$w N(-1, 0.9) + (1-w)N(1.2, 0.6)$ where $w = 0.35$. From the three 
distributions, 500 replicates of $K = 23$ quantiles were simulated for each of 
$n \in \{50, 150, 500, 1{,}000, 2{,}000, 5{,}000\}$. QM for each replicate of 
simulated quantiles was done using QGP, ORD, IND, SPL, and KDE. For the QGP, 
ORD, and IND models, $\theta = \{\theta_1, \theta_2, \theta_3, \theta_4\}$ 
where $\theta_c = (\mu_c, \sigma_c)$ for $c \in \{1,2,3,4\}$. The prior 
distributions assigned were $\mu_c \overset{ind}{\sim} N(5, 7^2)$,
$\sigma_c \overset{ind}{\sim} N(0, 6^2)\mathds{1}\{\sigma_c > 0\}$,
$w \sim Dir(\boldsymbol{\alpha}), \quad \alpha_c = 1$, and
$\nu \sim N(0, 3000^2)\mathds{1}\{\nu > 0\}$
% as below 
where $\nu = n$ for the QGP and ORD models 
and $\nu = 1/\sigma_{\rho}$ for the IND model. 
% (cite an example of using dirichlet prior ??). 
    % \begin{align*}
    %     \mu_c &\overset{ind}{\sim} N(5, 7^2) \\
    %     \sigma_c &\overset{ind}{\sim} N(0, 6^2)\mathds{1}\{\sigma_c > 0\} \\ 
    %     w &\sim Dir(\boldsymbol{\alpha}), \quad \alpha_c = 1 \\
    %     \nu &\sim N(0, 3000^2)\mathds{1}\{\nu > 0\}
    % \end{align*}
 %the problem with $\nu$ is that for the QGP and ORD the prior is on the variance whereas for IND it is on the sd.
Often when implementing posterior sampling for a mixture distribution model, 
one deals with an issue called the label-switching problem. This is where for a
mixture distribution parameter $\theta = \{\theta_1,...,\theta_C\}$, the model 
likelihood is the same for different permutations of $\theta$. This lack of 
identifiability for elements of $\theta$ makes parameter inference useless, 
but the predictive distribution may still be close to the true distribution 
\cite[]{stephens2000dealing}.
Because of this, the HMC posterior sampling chain was run longer than in the 
studies in the previous section. The chain was run for 80,000 steps where the 
first 20,000 draws were discarded as a burn-in. For this study, we were 
concerned only with QM and not with parameter inference, so the fact that the 
parameters are unidentifiable was not considered critical. However, when 
assessing the TV and KLD between the QM and true distributions, variational 
Bayes (VB) in \texttt{Stan} was used instead of HMC for fitting the models 
\cite[]{kucukelbir2015automatic}. This decision was made because of the lack of 
parameter identifiability and because the marginal parameter means are used to 
estimate the PDF. While MCMC methods perform better on parameter inference, 
VB methods have been shown to have predictive performance comparable to MCMC 
\cite[]{blei2017variational}.

@@

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.49\linewidth]{Images/quants_evd.png}
  % \caption{A subfigure}
  % \label{fig:sub1}
%\end{subfigure}%
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.49\linewidth]{Images/dens_evd.png}
  % \caption{A subfigure}
  % \label{fig:sub2}
%\end{subfigure}
\caption{QM fits of $K=23$ quantiles by KDE, SPL, IND, and QGP for 
$n \in \{50, 150, 500, 1{,}000\}$. The quantiles were sampled from the extreme 
value distribution $Ev(0,1)$. The quantile fits (left) show the true quantiles 
(black) with either the QM fit line (grey) or the credible intervals of 50\% 
(red) and 95\% (pink). 
The estimated PDF plots (right) show the true PDF (black) with either a the QM 
estimated PDF (grey) or the credible intervals of 50\% (red) and 95\% (pink).}
\label{fig:evd_fits}
\end{figure}

@@

%% If you have bib database file and want bibtex to generate the
%% bibitems, please use
%%
\bibliographystyle{elsarticle-harv}
\bibliography{master_bib}

>>>>

Along with the CDF and PDF, the QF is a defining function of a random variable. 
The QF, however, often receives less attention in standard statistical 
training than do the CDF and PDF \cite[]{parzen2004quantile}, and the CDF and 
PDF are used more often for modeling and inference.
% But the QF tends to be the little brother 
% \jarad{little brother?}
% of both the CDF and PDF in that the latter two functions are taught and used far more in statistical modeling and inference. 
% Nevertheless, the QF and several of its properties are central to the QGP model introduced in this paper. 
% \jarad{another throwaway sentence}
This section contains the definition of the QF, important properties, the 
definition of sample quantiles, and key central limit theorems (CLTs) of sample 
quantiles.

@@

The covariance function $\kappa_{\theta} (\cdot, \cdot)$ initially appears as 
if it will be difficult to evaluate and indeed can be, depending on the 
functional forms of $f_{\theta}$ and $Q_{\theta}$. But where these functions 
belong to distributions in a location-scale family, 
the covariance function can be greatly 
simplified. 
Noting that the covariance is a function of $f_{\theta}(Q_{\theta}(p)) = q_{\theta}(p)$, the QDF, if $\theta = (\mu, \sigma)$ where $\mu$ is a location parameter and $\sigma$ is a scale parameter, then $q_{\theta}(p) = \sigma q(p)$ where $q(p)$ is the QDF of a standard distribution having location 0 and shape 1. \cite{staudte2017shapes} calls this the location invariant and scale equivariant property. For location-scale distribution families, this greatly simplifies the form of $\kappa_{\theta}(\cdot,\cdot)$ so that it becomes
\[
    \kappa_{\theta}(p, p') = \sigma^{2}[p\wedge p' - p p'] q(p)q(p') 
\] 
Thus for estimating $\theta$ one only needs to estimate $\sigma^2$. A special 
case of the QGP for QM of a location-scale family is the normal QGP which we 
define and explore below.

@@

In section \ref{sec:simulation_analyses}, we analyze PIT transformed QGP model 
(\ref{eq:pit_qgp}) where $F_{\theta}$ is set as a normal mixture distribution. 
Fitting this model proved faster 
and produced better results than trying to fit the QGP model (\ref{eq:qgp}) and 
evaluating $Q_{\theta}$ the QF of a normal mixture. In fact we found that in 
cases where $F_{\theta}$ was a simpler function, such as a normal or 
exponential distribution, fitting model (\ref{eq:pit_qgp}) was at least as fast 
and the results were at least as good as fitting (\ref{eq:qgp}). Thus when 
possible, we elected to fit (\ref{eq:pit_qgp}).

@@

A number of metrics for measuring the distance between two continuous 
univariate distributions exist. Those we consider in this paper are the 
Wasserstein distance (WD) and a slight modification of it, the total variation 
(TV) or the statistical difference, and the Kullback-Leibler divergence (KLD). 
A brief overview of these and other metrics, some of their statistical uses 
and properties, and relationships between metrics is found in 
\cite{gibbs2002choosing}. These metrics were used in section 
\ref{sec:simulation_analyses} to assess how well different QM methods 
approximate a true distribution.

@@

We define a modified version of the WD, the uniform 1-WD (UWD1) in 
(\ref{eq:pit_wass_dist}). Here we take $F_X$ to be the CDF of a continuous 
random variable $X$. For a different continuous random variable $\xi$ with CDF 
$F_{\xi}$, $F_X(\xi)$ is a random variable with support on $[0,1]$. We let 
$F_{X,\xi}$ be the CDF of the random variable $F_X(\xi)$. Then the measure in 
(\ref{eq:pit_wass_dist}) measures how close $F_X(\xi)$ is the   CDF of 
a standard uniform distribution with CDF $\mathcal{U}$, 
noting that by the PIT, if 
$\xi \overset{D}{=} X$, $F_X(\xi)$ is the random variable of a standard uniform 
distribution. The integral in (\ref{eq:pit_wass_dist}) takes values between
0 and 1/2 \cite[]{645854}, and multiplying by two 
ensures the measure takes values between 0 and 1. This makes the UWD1 more
interpretable than the WD
where a value near 0 suggests $\xi$ and $X$ are ``near'' each other, and a value
near 1 suggests they are far apart.

@@

The top of figure \ref{fig:normal_cov_dists} shows the 90\% coverage of 
the posterior distribution credible intervals as $n$ increases for 
$\mu$, $\sigma$ and $n$ for five models. QGP-n and ORD-n models where $n$ is 
known are included along with QGP, ORD, and IND. 
The 90\% credible intervals of the 
parameters were calculated by computing the $0.05^{th}$ and $0.95^{th}$ 
quantiles from the posterior distribution samples. The coverage percentage is 
calculated as the percentage of the 500 replicates which the true parameter 
value was within the 90\% credible interval.
As expected, the nominal coverage for the QGP and ORD models is better than 
that of the IND model, and for the two models where $n$ is known the coverage 
is better than where it is unknown, particularly for lower values of 
$n$ and $K$. The coverage for the ORD models appears to be slightly better 
than for the QGP models though not by much, and the difference
decreases as $n$ increases. 
The superior performance of the ORD model is possibly because it
provides exact inference 
whereas the QGP model provides asymptotic inference.

@@

With the true distribution family being unknown, we made the modeling decision 
to set $F$ in the QGP model of (\ref{eq:pit_qgp}) to be a mixture of normals 
distribution taking the form of (\ref{eq:norm_mix}). This decision to use
a normal mixture was 
based on the common claim that any continuous distribution may be approximated 
by a finite mixture of normal distributions 
\cite[]{peel2000finite,nguyen2019approximations,nguyen2020approximation}. This 
claim may be optimistic, but we found that a four component normal mixture 
distribution can approximate the non-normally shaped distributions reasonably 
well. However, any function which meets the technical definition of a CDF would 
be appropriate to use for $F$, see \cite{gasthaus2019probabilistic} for an 
example.
After a small analysis comparing fits of one to six normal components,
the number of component 
distributions was chosen to be $C = 4$. The results, included in the 
supplementary materials, show that as the number of components in the mixture
distrubtion increases, the distribution approximation improves. But for the 
cases considered, the improvement from four to five components or more is small.

The three distributions from which data were simulated and sample quantiles 
estimated and to which QM methods were fit were the extreme value (EV) 
distribution with location 0 and scale 1 $EV(0,1)$, Laplace (La) with location 
0 and scale 1 $La(0,1)$, and a two component normal mixture (MIX) 
$w N(-1, 0.9) + (1-w)N(1.2, 0.6)$ where $w = 0.35$. From the three 
distributions, 500 replicates of $K = 23$ quantiles were simulated for each of 
$n \in \{50, 150, 500, 1{,}000, 2{,}000, 5{,}000\}$. QM for each replicate of 
simulated quantiles was done using QGP, ORD, IND, SPL, and KDE. For the QGP, 
ORD, and IND models, $\theta = \{\theta_1, \theta_2, \theta_3, \theta_4\}$ 
where $\theta_c = (\mu_c, \sigma_c)$ for $c \in \{1,2,3,4\}$. The prior 
distributions assigned were $\mu_c \overset{ind}{\sim} N(5, 7^2)$,
$\sigma_c \overset{ind}{\sim} N(0, 6^2)\mathds{1}\{\sigma_c > 0\}$,
$w \sim Dirichlet(\boldsymbol{1}_C)$, and
$\nu \sim N(0, 3000^2)\mathds{1}\{\nu > 0\}$
% as below 
where $\nu = n$ for the QGP and ORD models 
and $\nu = 1/\sigma_{\rho}$ for the IND model. 
% (cite an example of using dirichlet prior ??). 
    % \begin{align*}
    %     \mu_c &\overset{ind}{\sim} N(5, 7^2) \\
    %     \sigma_c &\overset{ind}{\sim} N(0, 6^2)\mathds{1}\{\sigma_c > 0\} \\ 
    %     w &\sim Dir(\boldsymbol{\alpha}), \quad \alpha_c = 1 \\
    %     \nu &\sim N(0, 3000^2)\mathds{1}\{\nu > 0\}
    % \end{align*}
 %the problem with $\nu$ is that for the QGP and ORD the prior is on the variance whereas for IND it is on the sd.
Often when implementing posterior sampling for a mixture distribution model, 
one deals with an issue called the label-switching problem. This is where for a
mixture distribution parameter $\theta = \{\theta_1,...,\theta_C\}$, the model 
likelihood is the same for different permutations of $\theta$. This lack of 
identifiability for elements of $\theta$ makes parameter inference useless, 
but the predictive distribution may still be close to the true distribution 
\cite[]{stephens2000dealing}.
Because of this, the HMC posterior sampling chain was run longer than in the 
studies in the previous section. The chain was run for 80,000 steps where the 
first 20,000 draws were discarded as a burn-in. For this study, we were 
concerned only with QM and not with parameter inference, so the fact that the 
parameters are unidentifiable was not considered critical. However, when 
assessing the TV and KLD between the QM and true distributions, variational 
Bayes (VB) in \texttt{Stan} was used instead of HMC for fitting the models 
\cite[]{kucukelbir2015automatic}. This decision was made because of the lack of 
parameter identifiability and because the marginal parameter means are used to 
estimate the PDF. While MCMC methods perform better on parameter inference, 
VB methods have been shown to have predictive performance comparable to MCMC 
\cite[]{blei2017variational}.

@@

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  % \includegraphics[width=\paperwidth]{Images/quants_dens_evd.png}
  \makebox[\textwidth][c]{%
        \adjustbox{trim=0 0 0 0,clip}{\includegraphics[width=1.3\textwidth]{Images/quants_dens_evd.png}}%
    }

\caption{QM fits of $K=23$ quantiles by KDE, SPL, IND, and QGP for 
$n \in \{50, 150, 500, 1{,}000\}$. The quantiles were sampled from the extreme 
value distribution $Ev(0,1)$. The quantile fits (left) show the true quantiles 
(black) with either the QM fit line (grey) or the credible intervals of 50\% 
(red) and 95\% (pink). 
The estimated PDF plots (right) show the true PDF (black) with either a the QM 
estimated PDF (grey) or the credible intervals of 50\% (red) and 95\% (pink).}
\label{fig:evd_fits}
\end{figure}

@@

%% If you have bib database file and want bibtex to generate the
%% bibitems, please use
%%
% \bibliographystyle{sn-apacite}
\bibliography{master_bib}

### 2025-10-08 supplemental.tex
manuscript edits

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1.3\linewidth]{Images/quants_dens_la.png}

@@

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=1.3\linewidth]{Images/quants_dens_gmix.png}
\caption{QM fits of $K=23$ quantiles by KDE, SPL, IND, and QGP for $n \in \{50, 150, 500, 1{,}000\}$. The quantiles were sampled from the two component normal mixture distribution $w N(-1, 0.9) + (1-w)N(1.2, .6)$ where $w = 0.35$. The quantile fits (left) show the true quantiles (black) with either the QM fit line (grey) or the credible intervals of 50\% (red) and 95\% (pink). 
The estimated PDF plots (right) show the true PDF (black) with either a the QM estimated PDF (grey) or the credible intervals of 50\% (red) and 95\% (pink). }
\label{fig:gmix_fits}
\end{figure}

>>>>

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \makebox[\textwidth][c]{%
        \adjustbox{trim=0 0 0 0,clip}{\includegraphics[width=1.3\textwidth]{Images/quants_dens_la.png}}%
    }

@@

\begin{figure}[hbt!]
\centering
%\begin{subfigure}{.5\textwidth}
  \centering
  \makebox[\textwidth][c]{%
        \adjustbox{trim=0 0 0 0,clip}{\includegraphics[width=1.3\textwidth]{Images/quants_dens_gmix.png}}%
    }
\caption{QM fits of $K=23$ quantiles by KDE, SPL, IND, and QGP for $n \in \{50, 150, 500, 1{,}000\}$. The quantiles were sampled from the two component normal mixture distribution $w N(-1, 0.9) + (1-w)N(1.2, .6)$ where $w = 0.35$. The quantile fits (left) show the true quantiles (black) with either the QM fit line (grey) or the credible intervals of 50\% (red) and 95\% (pink). 
The estimated PDF plots (right) show the true PDF (black) with either a the QM estimated PDF (grey) or the credible intervals of 50\% (red) and 95\% (pink). }
\label{fig:gmix_fits}
\end{figure}
